---
title: "Where Do People Drink The Most Beer, Wine And Spirits?"
author: "Sarah Jiang"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---


```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(fivethirtyeight)
library(here)
library(skimr)
library(janitor)
library(vroom)
library(tidyquant)
library(rvest)    # scrape websites
library(purrr)  
library(lubridate) #to handle dates
library(dplyr)
library(patchwork)
library(tidytext)
```



# Where Do People Drink The Most Beer, Wine And Spirits?


In this analysis, we are going to discuss where people drink the most beer, wine and spirits.

We we would like to install the package and load the original data in the beginning.

```{r, load_alcohol_data}
#Install the fivethirtyeight package  - as the data used i.e."drinks" is part of the package 

library(fivethirtyeight)
data(drinks)

```


Before we start, we want to answer these questions: What are the variable types? Any missing values we should worry about? 

```{r glimpse_skim_data}
#Summarize the data
glimpse(drinks)
skim(drinks)

```


*Answer:*The variable type for "country" column is character. The variable type for "beer_servings", "spirit_servings" and "wine_servings" is integer. The variable type for "total_litres_of_pure_alcohol" is double. 
There is no missing value in any of the columns.


Here, we make a plot that shows the top 25 beer consuming countries first.

```{r beer_plot}
#Select the top 25 beer consuming countries
beer_plot <- drinks %>%
  arrange(desc(beer_servings)) %>%
  select(country, beer_servings) %>% 
  head(25)

#Run the table
beer_plot

#Make the plot
ggplot(beer_plot,aes(x=reorder(country,beer_servings),y=beer_servings))+
  labs(title = "Top 25 Beer Consuming Countries", x="Country", y = "Beer Servings")+
  geom_bar(stat="identity")+
  coord_flip()

  

```

Next, we make a plot that shows the top 25 wine consuming countries

```{r wine_plot}

#Select the top 25 wine consuming countries
wine_plot <- drinks %>%
  arrange(desc(wine_servings)) %>%
  select(country, wine_servings) %>% 
  head(25)

#Run the table
wine_plot

#Make the plot
ggplot(wine_plot,aes(x=reorder(country,wine_servings),y=wine_servings))+
  labs(title = "Top 25 Wine Consuming Countries", x="Country", y = "Wine Servings")+
  geom_bar(stat="identity")+
  coord_flip()

```

Finally, we make a plot that shows the top 25 spirit consuming countries
```{r spirit_plot}
#Select the top 25 spirit consuming countries
spirit_plot <- drinks %>%
  arrange(desc(spirit_servings)) %>%
  select(country, spirit_servings) %>% 
  head(25)

#Run the table
spirit_plot

#Make the plot
ggplot(spirit_plot,aes(x=reorder(country,spirit_servings),y=spirit_servings))+
  labs(title = "Top 25 Spirit Consuming Countries", x="Country", y = "Spirit Servings")+
  geom_bar(stat="identity")+
  coord_flip()

```


> We can see that people from European countries love to drink wine and beer the most, as we can see lot of European countries' name on the first 2 graphs. On the other hand, people from Russian-speaking area love to drink spirits the most, as many of these countries appeared on the third graph.
We can also see that the popularity of each kind of alchohol is similar around the globe, as the average consumption of top 25 countris is similar. However, the differences of popularity among different countries seem to be the biggest in spirits and least in beer, as the standard deviation seems to be biggest in the spirit plot and least in the beer graph.



